Recent advances in random matrix theory applied to heavy-tailed multivariate time series

### 2015-09-07 papers/Number1/paper.tex
Revise the procedure for simulating eigenvalues.

\subsection{Overview of light-tailed
  case}\label{sec:light_tail_review}
As early as in 1962, T. W. Anderson \cite{anderson:1963} showed that,
when the entries of the data matrix has normal distribution ($\E
X_{ij} = 0$ for $i,j = 1,2,\dots, p$; $\E X_{ii}^2 = 2\sigma^2, \E
X_{ij}^2 = \sigma^2$ for $i \neq j$), the joint density function
of the eigenvalues of the sample covariance matrix is given by
\[
f(\lambda_1, \lambda_2, \dots, \lambda_p; \sigma, p) \propto \prod_{i=1}^p \exp\left(
  -{\lambda_i^2 \over 4\sigma^2}
\right) \prod_{i < j} (\lambda_i - \lambda_j)
\]
In 1988, Yin, Bai and Krishnaiah proved in
\cite{yin:bai:Krishnaiah:1988} the following result for a $p \times n$
matrix $\bf X$ with iid entries $X_{ij}$ under the assumption $\E
X_{11} = 0$, $\E X_{11}^4 < \infty$, and $q = \lim_{n \to \infty} p/n
> 0$
\[
\lambda_{\M}\left({1 \over n}\bf{XX'}\right) \overset{\as}{\to} \left(
  1 + \sqrt{q}
\right)^2 \E X_{11}^2 \text{ as }n \to \infty
\]
When $\E X_{11}^4 = \infty$, Yin and Silverstein proved in 1988
\cite{yin:silverstein:1988}
\[
\lim_{n \to \infty} \sup \lambda_{\M}\left({1 \over
    n}\bf{XX'}\right) = \infty
\]
Later in 2001, I. M. Johnstone showed that, when the data matrix has
standard normal distribution, i.e. $X_{ij} \sim N(0, 1)$, the largest
eigenvalue of the sample covariance matrix follows the Tracy-Widom
distribution:
\[
{ł_1 - \mu_{np} \over \sigma_{np}} \overset{d}{\to} W_1 \sim F_1
\text{ as }{n, p \to \infty}
\]
where $\mu_{np}$ and $\sigma_{np}$ are constants dependent on $p$ and
$n$ and the distribution function $F_1$ is given by
\[
F_1(s) = \exp\left\{
  -{1 \over 2} \int_{s}^\infty \left[
    q(x) + (x - s) q^2(x)
  \right] dx
\right\}
\]
where $q(x)$ is the unique solution to the Painlev\'e II differential
equation
\begin{eqnarray*}
  q''(x) &=& xq(x) + 2 q^3(x) \\
  q(x) &\sim& \text{Ai}(x) \text{ as } x \to \infty
\end{eqnarray*}
Figure \ref{fig:normal-TW} illustrates the density function of the largest
eigenvalue from a simulated sample and that of the Tracy-Widom.
\begin{figure}[htb!]
  \centering
  \includegraphics[scale=0.5]{normal-TW.pdf}
  \caption{Sample density function of the largest eigenvalue compared
    with the Tracy-Widom density function. The data matrix has
    dimension $200 \times 1000$. An ensamble of 2000 matrices are
    simulated.}
  \label{fig:normal-TW}
\end{figure}

>>>>

\subsection{Overview of light-tailed
  case}\label{sec:light_tail_review}
As early as in 1962, T. W. Anderson \cite{anderson:1963} showed that,
when the entries of the data matrix has normal distribution ($\E
X_{ij} = 0$ for $i,j = 1,2,\dots, p$; $\E X_{ii}^2 = 2\sigma^2, \E
X_{ij}^2 = \sigma^2$ for $i \neq j$), the joint density function
of the eigenvalues of the sample covariance matrix is given by
\[
f(\lambda_1, \lambda_2, \dots, \lambda_p; \sigma, p) \propto \prod_{i=1}^p \exp\left(
  -{\lambda_i^2 \over 4\sigma^2}
\right) \prod_{i < j} (\lambda_i - \lambda_j)
\]
In 1988, Yin, Bai and Krishnaiah proved in
\cite{yin:bai:Krishnaiah:1988} the following result for a $p \times n$
matrix $\bf X$ with iid entries $X_{ij}$ under the assumption $\E
X_{11} = 0$, $\E X_{11}^4 < \infty$, and $q = \lim_{n \to \infty} p/n
> 0$
\[
\lambda_{\M}\left({1 \over n}\bf{XX'}\right) \overset{\as}{\to} \left(
  1 + \sqrt{q}
\right)^2 \E X_{11}^2 \text{ as }n \to \infty
\]
When $\E X_{11}^4 = \infty$, Yin and Silverstein proved in 1988
\cite{yin:silverstein:1988}
\[
\lim_{n \to \infty} \sup \lambda_{\M}\left({1 \over
    n}\bf{XX'}\right) = \infty
\]
Later in 2001, I. M. Johnstone showed that, when the data matrix has
standard normal distribution, i.e. $X_{ij} \sim N(0, 1)$, the largest
eigenvalue of the sample covariance matrix follows the Tracy-Widom
distribution:
\[
{ł_1 - \mu_{np} \over \sigma_{np}} \overset{d}{\to} W_1 \sim F_1
\text{ as }{n, p \to \infty}
\]
where $\mu_{np}$ and $\sigma_{np}$ are constants dependent on $p$ and
$n$ and the distribution function $F_1$ is given by
\[
F_1(s) = \exp\left\{
  -{1 \over 2} \int_{s}^\infty \left[
    q(x) + (x - s) q^2(x)
  \right] dx
\right\}
\]
where $q(x)$ is the unique solution to the Painlev\'e II differential
equation
\begin{eqnarray*}
  q''(x) &=& xq(x) + 2 q^3(x) \\
  q(x) &\sim& \text{Ai}(x) \text{ as } x \to \infty
\end{eqnarray*}
Figure \ref{fig:3point-TW} illustrates the density function of the largest
eigenvalue from a simulated 3-point sample compared with that of the
Tracy-Widom.
\begin{figure}[htb!]
  \centering
  % \subfigure[Normal]{
  %   \includegraphics[scale=0.5]{normal-TW.pdf}
  % }
  \subfigure[3-point: $\P(X=sqrt(3)) = \P(X=-sqrt(3)) = 1/6$, $\P(X=0)=2/3$]{
    \includegraphics[scale=0.3]{3point-TW.pdf}
  }
  \caption{Sample density function of the largest eigenvalue compared
    with the Tracy-Widom density function. The data matrix has
    dimension $200 \times 1000$. An ensamble of 2000 matrices are
    simulated.}
  \label{fig:3point-TW}
\end{figure}
